Simulations of Ballot Polling Risk-Limiting Audits

### 2022-01-16 MultipleRoundSims/draft.tex
minor changes; size of latter figures

\begin{document}
\lstset{language=Python}
%
\title{Simulations of Ballot Polling Risk-Limiting Audits}
%add \thanks{} within the title brackets if want to refer to supporting organization ^^
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\begin{comment}
\author{Oliver Broadrick\inst{1} \and
Sarah Morin\inst{1} \and
Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
Filip Zag{\'o}rski\inst{3} }
%
\authorrunning{Broadrick, Morin et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Department of Computer Science, The George Washington University 
\email{obroadrick@gwmail.gwu.edu}\\
\and
Department of Computer Science, Stanford University 
\email{grantmcc@stanford.edu}\\
\and Wroclaw University of Science and Technology\\
\email{filip.zagorski@gmail.com}}
%
\end{comment}
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    In this paper we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of ballot polling risk-limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO, and End-of-Round (EoR) \BRAVO.    
    \BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
    In real audits, multiple ballots are drawn at a time, and \BRAVO is implemented as SO \BRAVO or EoR \BRAVO. 
    \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in a first round with stopping probability $0.9$ but requires a predetermined round schedule.
    %\BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. In real audits, election officials draw multiple ballots at a time, and \BRAVO rules may be implemented as SO \BRAVO or EoR \BRAVO, with neither being optimal for large first round sizes. \Minerva was recently proposed for use when the first-round stopping probability is large. It requires fewer first-round ballots, on average, than both SO \BRAVO and EoR \BRAVO for first round stopping probabilities of $0.9$ when the (true) underlying election is as announced. 
    It is an open question how these audits compare over multiple rounds and for lower stopping probabilities. 
    Our simulations use stopping probabilities of $0.9$ and $0.25$.
    %In this paper, we present results from simulations of multiple round audits with stopping probabilities of $0.9$ and $0.25$.
    %We examine both: first-round stopping probabilities of $0.9$, a common choice among election officials, and $0.25$, which would be more favorable to \BRAVO. 
    The results are consistent with predictions of the R2B2 open-source library for ballot polling audits. 
    %Because the \Minerva round schedule needs to be predetermined, we observe that the simulator is useful as a tool for understanding the behavior of \Minerva with the fixed round schedule. 
    We observe that both \BRAVO audits are more conservative than \Minerva, which stops with fewer ballots, for both first round stopping probabilities. 
    However, the advantage of using \Minerva decreases considerably for the smaller first round stopping probability, as one would expect.

@@

>>>>

\begin{document}
\lstset{language=Python}
%
\title{Simulations of Ballot Polling Risk-Limiting Audits}
%add \thanks{} within the title brackets if want to refer to supporting organization ^^
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Oliver Broadrick\inst{1} \and
Sarah Morin\inst{1} \and
Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
Filip Zag{\'o}rski\inst{3}\inst{4}}
%
\authorrunning{Broadrick, Morin et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Department of Computer Science, The George Washington University\thanks{Supported in part by NSF Award 2015253} 
\email{obroadrick@gwmail.gwu.edu}\\
\and
Department of Computer Science, Stanford University 
\email{grantmcc@stanford.edu}
\\
\and Wroclaw University of Science and Technology\thanks{Author was partially supported by Polish National Science Centre contract number DEC-2013/09/D/ST6/03927} and 
\and Votifica
\\
\email{filip.zagorski@gmail.com}
}
%

\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    In this paper we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of ballot polling risk-limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO, and End-of-Round (EoR) \BRAVO.    
    \BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
    In real audits, multiple ballots are drawn at a time, and \BRAVO is implemented as SO \BRAVO or EoR \BRAVO. 
    \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in a first round with stopping probability $0.9$ but requires a predetermined round schedule.
    %\BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. In real audits, election officials draw multiple ballots at a time, and \BRAVO rules may be implemented as SO \BRAVO or EoR \BRAVO, with neither being optimal for large first round sizes. \Minerva was recently proposed for use when the first-round stopping probability is large. It requires fewer first-round ballots, on average, than both SO \BRAVO and EoR \BRAVO for first round stopping probabilities of $0.9$ when the (true) underlying election is as announced. 
    It is an open question how these audits compare over multiple rounds and for lower stopping probabilities. 
    Our simulations use stopping probabilities of $0.9$ and $0.25$.
    %In this paper, we present results from simulations of multiple round audits with stopping probabilities of $0.9$ and $0.25$.
    %We examine both: first-round stopping probabilities of $0.9$, a common choice among election officials, and $0.25$, which would be more favorable to \BRAVO. 
    The results are consistent with predictions of the R2B2 open-source library for ballot polling audits. 
    %Because the \Minerva round schedule needs to be predetermined, we observe that the simulator is useful as a tool for understanding the behavior of \Minerva with the fixed round schedule. 
    We observe that both \BRAVO audits are more conservative than \Minerva, which stops with fewer ballots, for both first round stopping probabilities. 
    However, the advantage of using \Minerva decreases considerably for the smaller first round stopping probability, as one would expect.

@@

%\section{Acknowledgements}
%\label{sec:ack}
%This research was supported in part by NSF Awards 2015253
%and 1421373.

### 2022-01-16 MultipleRoundSims/sections/01-intro.tex
minor changes; size of latter figures

\Minerva was designed for R2 audits and applies its stopping rule once for each round. Thus it does not require the tracking of ballots that SO \BRAVO does. Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit\footnote{Their proof assumes that the number of ballots drawn in each round is know beforehand} and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds, the two audits have the same pre-determined (before any ballots are drawn) round schedule and the underlying election is as announced. They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for first round sizes with a large probability of stopping when the (true) underlying election is as announced.

@@

\subsection{Our Contributions}
For a two candidate plurality contest with a risk limit of $10\%$, we show the following
\begin{enumerate}
\item Even when the first round stopping probability is as small as $0.25$, the number of ballots required for \Minerva is smaller than that required by SO \BRAVO and EoR \BRAVO. However, the improvement is considerably smaller than that when the stopping probability is $0.9$. 
\begin{itemize}
\item The number of ballots required by SO \BRAVO for a first round stopping probability of $0.9$ is about a third more than that required by \Minerva. On the other hand, for a first round stopping probability of $0.25$, it requires only about a tenth more ballots than does \Minerva.  
\item The number of ballots required by EoR \BRAVO for a first round stopping probability of $0.9$ is about twice those required by \Minerva. On the other hand, for a first round stopping probability of $0.25$, it requires only about a fourth to a half more ballots (depending on margin) than does \Minerva.  
\end{itemize}
\item For a first round stopping probability of $0.9$, when consequent \Minerva rounds are the same size (multiplying factor $1$), consequent conditional stopping probabilities are about $0.75$ and $0.74$ respectively for rounds two and three. When the multiplying factor is $1.5$, the conditional stopping probabilities for rounds two and three are $0.91$ and $0.83$ respectively. Both our simulator and the code estimating probabilities and round sizes are flexible enough to enable the study of various predetermined round schedules. 
\end{enumerate}

>>>>

\Minerva was designed for R2 audits and applies its stopping rule once for each round. Thus it does not require the tracking of ballots that SO \BRAVO does. Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds, the two audits have the same pre-determined (before any ballots are drawn) round schedule and the underlying election is as announced\footnote{Their proof assumes that the number of relevant ballots drawn in each round is know beforehand. In MINERVA, the number of ballots drawn in each round is determined before any ballots are drawn. Because invalid ballots and ballots that are inconsequential for the contest being audited would be drawn in addition to relevant ballots, the assumption used by the proof is not true in general. (We are grateful to Philip Stark for drawing our attention to this.) However, any variation in number of relevant ballots drawn for a fixed round size would be random and not chosen by an adversary; the proof showing the risk-limiting property of MINERVA could hence be extended.} . They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for first round sizes with a large probability of stopping when the (true) underlying election is as announced.

@@

\subsection{Our Contributions}
For a two candidate plurality contest with a risk limit of $10\%$, we observe the following about the total number of ballots drawn over five rounds:
\begin{enumerate}
\item Even when the first round stopping probability is as small as $0.25$, the number of ballots required for \Minerva is smaller than that required by SO \BRAVO and EoR \BRAVO. However, the improvement is considerably smaller than that when the stopping probability is $0.9$. 
\begin{itemize}
\item The number of ballots required by SO \BRAVO for a first round stopping probability of $0.9$ is about a third more than that required by \Minerva. On the other hand, for a first round stopping probability of $0.25$, it requires only about a tenth more ballots than does \Minerva.  
\item The number of ballots required by EoR \BRAVO for a first round stopping probability of $0.9$ is about twice those required by \Minerva. On the other hand, for a first round stopping probability of $0.25$, it requires only about a fourth to a half more ballots (depending on margin) than does \Minerva.  
\end{itemize}
\item For a first round stopping probability of $0.9$, when consequent \Minerva rounds are the same size (multiplying factor $1$), consequent conditional stopping probabilities are about $0.75$ and $0.74$ respectively for rounds two and three. When the multiplying factor is $1.5$, the conditional stopping probabilities for rounds two and three are $0.91$ and $0.83$ respectively. Both our simulator and the code estimating probabilities and round sizes are flexible enough to enable the study of various predetermined round schedules. 
\end{enumerate}

### 2022-01-16 MultipleRoundSims/sections/04-expts.tex
minor changes; size of latter figures

All of the audits discussed in this paper are modeled as binary hypothesis tests. Under the alternative hypothesis, $H_a$, the announced outcome is correct. In particular, the true underlying ballot distribution is given by the announced ballot tallies. Under the null hypothesis, $H_0$, the true outcome is a tie (or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the loser).
The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie. (Vora shows that this is the maximum risk \cite{Bayesian-RLA}.) 
\begin{definition}[Risk]
The maximum risk $R$ of audit $\mathcal{A}$ with sample $X\in \{0,1\}^*$ drawn from 
the true underlying distribution of ballots is
$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0].$
\end{definition}

This leads us to the following definition of an $\alpha$-RLA.
\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
An audit $\mathcal{A}$ is a Risk Limiting Audit with 
risk limit $\alpha$ iff 
$R(\mathcal{A}) \le \alpha.$
\end{definition}
Note that an audit $\mathcal{A}$ includes all audit parameters (risk limit, round sizes, etc.).

We present measures of stopping probability in the $j^{th}$ round of the audit, given that the underlying election is as announced.
\begin{definition}[Stopping Probability]
The stopping probability $S_j$ of an audit $\mathcal{A}$ in round $j$ is 
$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
\end{definition}
Experimentally, using our simulations, $S_j$ would be estimated by the fraction of audits that stop in round $j$. Note that $\sum _j S_j(\mathcal{A})=1$. We can also consider the cumulative stopping probability: 
\begin{definition}[Cumulative Stopping Probability]
The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is 
$$C_j(\mathcal{A})= \sum_{i=1}^j S_j$$
\end{definition}
Experimentally, using our simulations, $C_j$ would be estimated by the fraction of audits that stop in or before round $j$.

>>>>

All of the audits discussed in this paper are modeled as binary hypothesis tests. Under the alternative hypothesis, $H_a$, the announced outcome is correct. In particular, the true underlying ballot distribution is given by the announced ballot tallies. Under the null hypothesis, $H_0$, a tie is the correct outcome \footnote{or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the loser}.
The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie \cite{Bayesian-RLA}. Note that an audit $\mathcal{A}$ includes all audit parameters (maximum risk, round sizes, etc.).

\begin{definition}[Risk]
The maximum risk $R$ of audit $\mathcal{A}$ with sample $X\in \{0,1\}^*$ drawn from 
the true underlying distribution of ballots is
$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0].$
\end{definition}

This leads us to the following definition of an $\alpha$-RLA.
\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
An audit $\mathcal{A}$ is a Risk Limiting Audit with 
risk limit $\alpha$ iff 
$R(\mathcal{A}) \le \alpha.$
\end{definition}

We present measures of stopping probability in the $j^{th}$ round of the audit, given that the underlying election is as announced.
\begin{definition}[Stopping Probability]
The stopping probability $S_j$ of an audit $\mathcal{A}$ in round $j$ is 
$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
\end{definition}
Experimentally, using our simulations, $S_j$ would be estimated by the fraction of audits that stop in round $j$. Note that $\sum _j S_j(\mathcal{A})=1$. We can also consider the cumulative stopping probability: 
\begin{definition}[Cumulative Stopping Probability]
The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is $C_j(\mathcal{A})= \sum_{i=1}^j S_j$
\end{definition}
Experimentally, using our simulations, $C_j$ would be estimated by the fraction of audits that stop in or before round $j$.
